---
title: "gemeinden weekly"
output: html_document
date: '2022-06-03'
---


This file creates weekly interpolated population data for communities (Gemeinden).

Setting `GKZ_to_use` fo a vector of Gemeindekennzahlen (within this file or in an other filesx purling this file) will produce the result only for the selected Gemeinden.

If `GKZ_to_use` is NULL, a complete dataset for all Gemeinden will be produced.       
If `GKZ_to_use` is undefined, it will be set to NULL.          
This takes time and memory. I set        
`R_MAX_VSIZE=128Gb` on a machine with 64Gb of real memory, and it took about 22 minutes.




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(fs)
  library(here)
  library(lubridate)
  library(readxl)
  library(furrr)
  library(purrr)
  library(numbers)
  library(tictoc)
})
```

```{r}
source(path("utils","purl_and_source.R"))
```

```{r}
purl_and_source(path("utils","date_utils.Rmd"))
```

```{r}
save_path <- path(here(),"r_data")
```

```{r}
purl_and_source(path(here(),"utils","get_region_names.Rmd"))
```


```{r}
load(path(save_path,"pop_GKZ_1y_gender_yearly.RData"))
```

```{r}
pop_GKZ_1y_gender_yearly |> str()
```

```{}
dates_years_weeks <-
  tibble(
    Datum=seq(as.Date("2001-12-31"),as.Date("2025-01-01"),by="7 days"),
    Jahr=isoyear(Datum),
    Woche=isoweek(Datum)
  ) |>
  group_by(Jahr) |>
  mutate(total_weeks=max(Woche)) |>
  ungroup() |>
  mutate(Jahr_Woche=Jahr+(Woche-1)/total_weeks) |>
  select(-total_weeks) |>
  mutate_at(vars(Jahr,Woche),as.integer)
```


```{r}
#if(is.null(GKZ_to_use)){
  data_to_use <- pop_GKZ_1y_gender_yearly
#} else {
#  data_to_use <- pop_GKZ_1y_gender_yearly |>
#    filter(GKZ %in% GKZ_to_use)
#}
#rm(pop_GKZ_1y_gender_yearly)
```



```{r}
tic()
expand_grid(
  datum_jahr_woche() |> select(-Quartal,-Wochen_im_Quartal),
  GKZ = data_to_use |>
    pull(GKZ) |> 
    unique() |>
    as.integer() ,
  Alter= 0:100 |> as.integer(), 
  Geschlecht=c("m","w")) |> 
  left_join(data_to_use  |>
              select(-Name) |>
              mutate(Woche=1L)) |>
  mutate(pop = pop |> as.numeric()) ->
  data_incomplete
rm(data_to_use)
gc() |> invisible()
gc() |> invisible()
toc()
data_incomplete |>
  str()
```


```{r}
save(data_incomplete, 
     file=path(save_path,"data_incomplete.RData"))
gc() |> invisible()
```

```{r}
tic()
file_size(path(save_path,"data_incomplete.RData"))
toc()
```

-=-=-=-=-=-=-=-=-=-=

Here the loop starts

```{r}
plan(multisession, workers = 8)

bl_num <- 1:9
pop_GKZ_1y_gender_weekly <-
  tibble()


#bl_num <- 1
tic()
for (bl in bl_num) {
  load(path(save_path, "data_incomplete.RData"))
  data_incomplete |>
    filter((floor(GKZ / 10000) |> as.integer()) == bl) ->
    filtered_df
  if(nrow(filtered_df) == 0) next
    filtered_df |>  
    split(~ GKZ + Alter + Geschlecht) -> xxx

  rm(data_incomplete)
  gc() |> invisible()
  gc() |> invisible()

  xxx |>
    future_map_dfr(
      \(x)
        mutate(x,
          pop_new = approxfun(
            x |>
              filter(!is.na(pop)) |>
              pull(Jahr_Woche),
            x |> filter(!is.na(pop)) |>
              pull(pop),
            rule = 2
          )(Jahr_Woche)
        ) |>
          select(-pop) |>
          rename(pop = pop_new),
        .progress=TRUE
    ) ->
  yyy
  gc() |> invisible()
  gc() |> invisible()
  bind_rows(
    pop_GKZ_1y_gender_weekly,
    yyy
  ) ->
  pop_GKZ_1y_gender_weekly
  rm(yyy)
  gc() |> invisible()
  gc() |> invisible()
}
toc()
```
```{r}
pop_GKZ_1y_gender_weekly |> str()
```

```{r}
unlink(path(save_path,"data_incomplete.RData"))
```
```{r}
pop_GKZ_1y_gender_weekly |> 
  select(-Jahr_Woche) ->
  pop_GKZ_1y_gender_weekly  
```

```{r}
tic()
#if(is_null(GKZ_to_use)){
  save(pop_GKZ_1y_gender_weekly,
       file=path(save_path,"pop_GKZ_1y_gender_weekly.RData"))
#} else {
#  df_name <- paste(c("pop_GKZ_1y_gender_weekly",GKZ_to_use),collapse="_")
#  assign(df_name,pop_GKZ_1y_gender_weekly,
#         envir=.GlobalEnv)
#  save(list=df_name,
#       file=path(save_path,
#                 paste0(df_name,".RData")))
#}
toc()  
```


